---
title: "03 Vectors"
output: html_document
date: '2022-08-07'
editor_options: 
  chunk_output_type: console
---

## 3.1 Introduction

Two flavours: atomic vectors and lists
Atomic: all elements must have the same type
NULL: generic zero length vector

Vector attributes
- *dimension* turns vectors into matrices and arrays
- *class* powers the S3 object system

## 3.2 Atomic vectors

Four primary types: logical, integer, double and character
Two rare types: complex and raw

### 3.2.1 Scalars

Each of the four primary types has a *scalar*: a special syntax to create an individual value
- Logicals (`TRUE` or `FALSE`)
- Doubles: decimal (`0.1234`), scientific (`1.234e4`) or hexadecimal (`0xcafe`). Special values are `-Inf`, `Inf` and `NaN`
- Integer: must be followed by `L` (`1234L`)
- Strings surrounded by `"` or `'`

### 3.2.2 Making longer vectors with `c()`

Use `c()`, short for combine
```{r making longer vectors with c()}
lgl_var <- c(TRUE, FALSE)
int_var <- c(1L, 6L, 10L)
dbl_var <- c(1, 2.5, 4.5)
chr_var <- c("these are", "some strings")
```

When the inputs are atomic vectors the output is an atomic vector
```{r inputs = atomic, output = atomic}
c(c(1, 2), c(3, 4))
```
Determine the type of a vector with `typeof()` and length with `length()`
```{r determine the type and length of a vector}
typeof(lgl_var)
typeof(int_var)
typeof(dbl_var)
typeof(chr_var)

length(lgl_var)
```

### 3.2.3 Missing values

Missing values tend to be infectious

```{r missing values tend to be infectious}
NA > 5
10 * NA
!NA
```

There are a few exceptions

```{r missing value exceptions}
NA ^ 0
NA | TRUE
NA & FALSE
```

A common mistake when determining if values in a vector are missing
```{r common mistake when determining missing values in a vector}
x <- c(NA, 5, NA, 10)
x == NA

# instead use is.na()
is.na(x)
```

### 3.2.4 Testing and coercion

You can *test* if a function is a given type using `is.*()` functions

When you attempt to combine different vector types they will be *coerced* in a fixed order: character → double → integer → logical

```{r combining a character and an integer yields a character}
str(c("a", 1))
```

Coercion often happens automatically
```{r coercion often happens automatically}
x <- c(FALSE, FALSE, TRUE)
as.numeric(x)

# Total number of TRUEs
sum(x)

# Proportion that are TRUE
mean(x)
```

Generally you can deliberately coerce using an `as.*()` function
```{r deliberate coerce using as.integer()}
as.integer(c("1", "1.5", "a"))
```

### 3.2.5 Exercises

1. How do you create raw and complex scalars? (See `?raw` and `?complex`.)
```{r 3.2.5 Exercise 1}
# raw() creates a raw vector of specified length
raw(3)

# as.raw() attempts to coerce an argument to be of raw type
as.raw(10)

# complex() creates a complex vectors
complex(3)  # specified length generates a vector of complex zeroes
complex(real = 2, imaginary = 0.5)
complex(modulus = sqrt(8), argument = pi / 4)

# as.complex() attempts to coerce an argument to be of complex type
as.complex(10)
```


2. Test your knowledge of the vector coercion rules by predicting the output of the following uses of c():

`c(1, FALSE)`
`c("a", 1)`
`c(TRUE, 1L)`

3. Why is `1 == "1"` true? Why is `-1 < FALSE` true? Why is `"one" < 2` false?

4. Why is the default missing value, `NA`, a logical vector? What’s special about logical vectors? (Hint: think about `c(FALSE, NA_character_)`.)

5. Precisely what do `is.atomic()`, `is.numeric()`, and `is.vector()` test for?

## 3.3 Attributes

Data structures like matrices, arrays, factors and date-time are built on top of atomic vectors by adding attributes

### 3.3.1 Getting and setting

### 3.3.2 Names

### 3.3.3 Dimensions

### 3.3.4 Exercises

1. How is setNames() implemented? How is unname() implemented? Read the source code.

2. What does dim() return when applied to a 1-dimensional vector? When might you use NROW() or NCOL()?

3. How would you describe the following three objects? What makes them different from 1:5?

```{r}
x1 <- array(1:5, c(1, 1, 5))
x2 <- array(1:5, c(1, 5, 1))
x3 <- array(1:5, c(5, 1, 1))Copy
```

4. An early draft used this code to illustrate structure():

```{r}
structure(1:5, comment = "my attribute")
```

But when you print that object you don’t see the comment attribute. Why? Is the attribute missing, or is there something else special about it? (Hint: try using help.)